% A SHORT INTRODUCTION
% Just for fun, maybe do one sentence per line.
% Just give an eloquent and broad explanation of the problem that I'm addressing and why I'm doing it, then give a somewhat detailed chapter summary. Follow the model of the abstract for like 1500 words and it will be fine.
% Is the goal analysis of these new musics? Or is the goal better understanding of NOTATION at large? Be extremely lucid here.

%\begin{notestuff}
%Elaborate the problem.
%\end{notestuff}
\addcontentsline{toc}{chapter}{\protect\numberline{}Introduction}

%Western music notation is overencumbered.
%As the \textit{lingua franca} for so many disparate musical practices, it is pressed into service in a number of different capacities---many of them conflicting.
%For Western classical performers, it must serve as a recipe to be more-or-less strictly obeyed such that the resulting sounds might conform with appropriate precision to the composer's (real or imagined) wishes.
%For the analyst, notation must be able to stand in well-defined, logically consistent relationships such that patterns may be observed, predictions made, and conclusions drawn.
%For the transcriber, it must map one-to-one with observed sonic phenomena; able to represent minute sonic details well enough to be fit for archiving or later reproduction.
%Finally, for the improviser, notation must serve to induce creative action, providing a material springboard for musical creativity while simultaneously delineating the boundaries of an authentic performance.
Western music notation is overencumbered. Having served as a \textit{lingua franca} for so many musical practices over the last millennium, it has been pressed into service in a great many capacities which often functionally conflict with one another.
For the ``classical'' musician, notation must behave as a recipe to be more-or-less strictly obeyed according to a composer's (real or imagined) wishes. 
For the analyst, it must stand in logical, well-structured relationships that remain stable from piece to piece such that patterns may be observed and conclusions drawn.
For the transcriber, it must map one-to-one with observed sonic phenomena; able to represent minute sonic details well enough to be fit for archiving or reproduction.
Conversely, for the improviser it must remain usefully vague, serving to induce musical invention by providing a material springboard for creativity.

Given these requirements, Western notation has proven admirably metastatic---able to be augmented, pared down, or redefined according to the needs of its adoptive community of practice.
When medieval scholar-clergy needed a new way of transmitting and archiving sacred melodies, notation was altered such that it might serve as a fixed point of reference, supplanting earlier, vaguer neumes.
When improvised accompaniment became \textit{de rigueur} for Baroque soloists, the music's composers, arrangers, and engravers developed new symbols to clarify music's harmonic underpinnings.
Then, when improvisation subsequently fell out of vogue, new symbols allowed players to more closely track composers' pre-conceived sound-concepts.

%Notation's malleability held as much for new music in the twentieth century as it did for that of the eighteenth. 
%Owing to a variety of factors, the mid-to-late twentieth century saw a veritable explosion of novel notation techniques; some quite durable but many which served only a single piece.
%A significant proportion of these new notations (perhaps the majority) were developed in response to composers' growing desire for sonic indeterminacy---for scores which would generate predictably or unpredictably unique listening experiences each time they were realized. 
%In a way this was the same as any earlier adaptation---artists saw a new need and developed technology to fill it.
%There was a kernel of difference, however: during this period notation \textit{itself} became a vector for artistry and expression to an extent theretofore unknown.
%Composers, it seems, became increasingly interested in notation's power to effect not only a resultant sound-world, but new relationships between artist and interpreter: thus they began to compose notation as well as music.
%As a result, many of these new notations prove to be fertile ground for new scholarship; not only on their own merits but because they have the potential to provide insights into the practice of notating music more generally. 

This notational malleability held as much for the bleeding-edge modernism of the twentieth century as it had in earlier eras.
The 1950s and 1960s exploded with novel notation schemes---some short lived and some more durable.
Though their invention and implementation were motivated by a number of factors, the majority of new notations, it seems, arose from composers' desire for some form of sonic indeterminacy---for scores which would yield unique listening experiences each time they were realized.
In a sense, this was the same as any earlier innovation: artists discovered a need and developed new technology to fill it. The kernel of difference lay in the unprecedented extent to which the notation \textit{itself} became a vector for artistic expression.
Composers, it seems, had grown increasingly concerned with its power not only to realize a particular sound-world, but to mediate player/author/score interactions as well.
As a result, these twentieth-century notations prove fertile ground for new scholarship---not only on their own merits, but because of their potential to grant insight into literate music-making at large.

As a composer-improviser with an atypically tortuous relationship to jazz and contemporary Western concert music, I took particular interest in the notation practices situated nearest the juncture between these two fields, those of the AACM (particularly those of Anthony Braxton, my old mentor Roscoe Mitchell, Wadada Leo Smith, and George Lewis). 
Familiarizing myself with these composers, in turn, necessarily led to greater awareness of their European and American ``concert music'' peers and forebears: artists in the New York School (John Cage, Earle Brown, Christian Wolff, Morton Feldman) and the titans of Euro-modernism (Karlheinz Stockhausen, Gy\"orgy Ligeti, Helmut Lachenmann, Horațiu R\u{a}dulescu, to name a few.)

As the purview of my research expanded, so, too, did my interest not only in these enigmatic symbols themselves, but in the motivations that lay behind their deployment
Some important questions arose: What factors do composers consider when designing new notations?
How does one use notation to balance cogent musical form with an ever-changing sonic surface?
What differentiates notations which properly \textit{encode} from those which seek only to effect ekphrasis, a necessarily imperfect translation from one aesthetic domain to another? 

%\begin{notestuff}
%    Problematize the literature.
%\end{notestuff}
Of course, there is no shortage of literature attempting to clarify and contextualize these new practices. 
Scholars (since at least Umberto Eco) have recognized the novelty and significance of these fundamentally ``incomplete'' scores which require players' creative contributions in order to be finalized in performance. 
However, one finds that much of this literature lacks focus and precision. 
Far more attention seems to be paid to new notations' aesthesis and physical trace than is paid to their function.
Many writers attempt to discuss novel notations without a robust notion of what differentiates the new from the old, and without a unifying narrative able to account for the ways openness has \emph{always} been a part of notation's function and use.
Further, writers routinely conflate composers' underlying philosophical commitments with the tools they use to inscribe their music.
Thus categories of notation are unnecessarily multiplied: ``aleatoric notation,'' ``indeterminate notation,'' ``improvisatory notation,'' and worst of all, the ultra-vague ``graphic notation,'' all of which lack unambiguous definitions that would distinguish them from earlier forms and from one another. 
As I see it, these issues point to a marked lack of conceptual rigidity and descriptive language when it comes to notation's use and function, particularly as regards complex, late-century ``open'' notation schemes. 
    
This dissertation is thus an attempt to address, in part, lacunae in the study of these ``open notations,'' i.e., of notations oriented toward finer mediation of sonic indeterminacy.
To be more specific, my interest principally lies in notations which center syntactically and semantically well-defined symbols, as these seem to frequently take a backseat (in practice and in the literature) to more ``radical'' methods which deliberately strip away notation's syntax and semantics, granting the performer so much agency that the composer's contributions become practically negligible.\footnote{...though to be clear these ``asemantic'' scores, too, deserve greater scrutiny.}
I contend that systems which instead \textit{build upon} the strengths of traditional notation, deliberately amplifying its ability to mediate composer/performer relationships have the most to tell us about the way we interact with the score.
Thus, over the course of this document I explore a number of these systems via a multi-pronged approach.
%\begin{notestuff}
%    Chapter breakdown.
%\end{notestuff}
%\paragraph{Chapter One} % Remove these \paragraphs before go time.

Chapter One conducts a rapid historical survey, developing a narrative which locates notational fixity and openness as core sites of innovation in Western art music notation. 
Beginning with Guido d'Arezzo and ending in the 1960s, I portray the history of Western notation as a gradual ebb and flow in the degree of coupling between the printed page and its sonic products. 
Specifically, I note the way that the form and function of notation have always responded to the particular needs of its user-base, though, crucially, that form and function are not always neatly tied together.
Where significant advances in either occur, I examine specific instances where this change is made manifest, which range from the development of \textit{cantare super librum} (which I take to be one of the earliest modern literate/improvisatory practices), to the decline of notation's ``openness'' in the late-eighteenth and nineteenth centuries, to the Afro-diasporic return to notation for improvisers in the mid-twentieth.
In addition, I briefly examine some of the first (predominantly New York School) forays into the new notations at the heart of this assessment. 
Here I render notational openness and fixity as first emergent and later deliberately exploited musical parameters, always already linked to the very notion of notated music.

%\paragraph{Chapter Two}
Chapter Two uses concepts explored in the first chapter to examine several aspects of notation's function in detail, ultimately taking steps toward a more robust typology of (open) notations
Beginning from first principles, I pose a new model by which we might conceive of notation's semantic content---i.e., what these symbols represent and how they convey their message.
So as to facilitate the development of a more useful critical lexicon, I challenge what I take to be the prevailing (often insufficiently-articulated) ``folk-semiosis'' pertaining to the function of music notation generally.
Following this, I examine essays by three authors, Umberto Eco, Pierre Boulez, and Gy\"{o}rgy Ligeti (all roughly contemporary with the mid-century notation boom), who each seek to address the roles and goals of various open notations. 
In particular, I spend significant time on Ligeti's 1965 „Neue Notation: Kommunikationsmittel oder Selbstzweck?,” a woefully under-appreciated essay in which Ligeti, a scholar-composer situated at the bleeding edge of Western concert music innovation, lends unique perspective to questions of notation's function.
Most importantly, Ligeti examines a number of contemporary neonotational works and, in so doing, puts forward a new typology so as to accommodate these new pieces.
Finally, based on these insights I propose two new descriptors, ``traversal'' and ``hybridity,'' used to describe the ways composers combine and move through notations at varying degrees of fixity and with varying levels of semanticity. 

%\paragraph{Chapter Three}
Chapter Three demonstrates an application of this new descriptive paradigm, examining two ``work complexes'' by late-century composers Anthony Braxton and Horațiu R\u{a}dulescu through the lens of a modified Ligetian typology.
Braxton and R\u{a}dulescu, two innovators belonging to very different communities of musical practice, each employ complex, bespoke notation schemes which, to speak broadly, deliberately parametrize notational fixity and openness in various ways.
Here, I attempt to lead by example, conducting a notation-centric analysis comparing the two composers' systems on a symbol-for-symbol basis and describing what I take to be the most salient points of overlap and departure between them.
Through this analysis, I draw conclusions pertaining not only to Braxton's and R\u{a}dulescu's writing methods, but also to their underlying philosophical commitments and to the unexamined potential of well-defined ``open'' notations generally.

%\paragraph{Chapter Four}

Finally, Chapter Four serves to document the author's yearslong creative investigation of the topics explored in previous chapters; specifically, the design and implementation of \{O-G\} (``Otto-Glyphs'') a novel notation scheme for improvising musicians.
I begin by discussing the system's initial catalysts: dissatisfaction with some of my formative experiences with scored improvisation and my subsequent exposure to a number of artists who had succeeded in creatively circumventing these same problems.
I continue by summarizing \{O-G\}'s nascence, addressing a number of important milestones including the first piece to be formally composed in the scheme as well as the formal ``instruction manual'' which eventually became the core tool by which I inducted new players into the system.
The system (and my pedagogical method) would ultimately be put to the test in \textit{I Die Each Time I Hear the Sound}, a concert of original works composed either entirely in \{O-G\} or with well-integrated \{O-G\}/traditional notation.
I thus conclude by summarizing several of the more noteworthy compositions from this series before conducting a broad-level assessment of \{O-G\}'s successes and failures according to the criteria delineated at the project's outset.



